%!TEX root = umthsmpl.tex
\chapter{Overview of Blockchain Systems}

\section{Basic Operation}

\para{Account balances.} A Bitcoin is a unit of currency, which is
fungible, divisible (up to eight decimal places), and recombinable.
It is measured as a balance across multiple accounts, which are
themselves manifested in \emph{addresses}.$\!$\footnote{Internally,
  Bitcoins exist only as ``unspent transaction outputs'' (UTXO), but
  users of the system think of them as balances in addresses, and that
  view does not affect the results of this paper.  } Each address
comprises a stored asymmetric cryptographic key and an associated
balance of Bitcoin. The public portions of an address are the public
key and the balance of coin. When an address is involved in a
\emph{transaction} with one or more other addresses, Bitcoins are
transferred among them.
%Addresses are explicitly pseudonyms, and not tied to a particular individual.
% ; further, empty addresses can be created at no cost beyond
% generating an asymmetric key pair.

\para{Roles.} Users wishing to exchange coins broadcast the details of
their transactions over Bitcoin's p2p network, signed with their
private keys. A set of \emph{miners} on the p2p network verify that
each transaction is signed correctly and does not conflict with
another transaction. Miners independently agglomerate a set of valid
transactions into a \emph{block} and attempt to solve a predefined
proof-of-work (POW) problem involving this block and a chain of prior
valid blocks. In Bitcoin, the POW computation is dynamically
calibrated to take approximately ten minutes per block. The first
miner to solve the problem broadcasts his solution to the network,
adding it to the ever-growing \emph{blockchain}; the miners then start
over, with the appended blockchain and the set of transactions that
were not added as part of the previous block. When transactions appear
in a block, they are considered \emph{confirmed}, and each subsequent
block provides additional confirmation. The miners' incentive for
discovering a block is a reward of coins, called the {\em coinbase},
consisting of a predetermined {\em block reward} (currently 12.5 BTC)
and fees from transactions included in the block.

\emph{Full nodes} are peers in the network that do not mine, but do
generate, validate, and propagate transactions and blocks to other
nodes including miners. Consumers (i.e., those who purchase goods or
services) typically have no need to process and validate all
transactions, so they can instead operate \emph{simple payment
  verification} (SPV) nodes that process, store, and transmit data
involving only addresses-of-interest, which are typically addresses
they control, make payments to, or receive payments from. SPV nodes
rely on full nodes to relay transactions-of-interest.

% These SPV nodes are intended to run on resource-constrained devices,
% such as smartphones, which typically possess between 16--64 GB of
% storage shared among all applications. They are the leaves of the
% Bitcoin p2p network (i.e., they connect to only a few full nodes and
% never other SPV nodes), relying on full nodes to operate as a
% liaison between them and the miners. To remain anonymous to full
% nodes, addresses-of-interest are deposited by the SPV node into a
% Bloom filter. Bloom filters are a specially prepared array that is
% an efficient digest of the elements in a set. With the array,
% another peer can determine, with high probability, whether an
% element is contained in the set that produced the array. This filter
% is passed to neighboring full nodes that continuously monitor all
% transactions, identify transactions-of-interest using it, and
% forward those transactions to the SPV node.

\para{Bitcoin transaction consistency.} 
The main goal of the Bitcoin p2p network is to provide a consistent
view of blocks and unconfirmed transactions across all network peers.
Each peer maintains a local snapshot of the transactions in a memory
pool dubbed the \emph{mempool}. Blocks consist of a list of
transactions that have already (almost always) been broadcast to
miners and full nodes in the network.

To announce a new block, a miner lists all transactions contained in
the new block along with a header that provides an easily verifiable
\emph{proof-of-work} (POW) solution.  When a full node or miner receives a new block, it
validates each transaction in the block and the proof of work.

Due to propagation delays in the network, it is possible for the
miners to receive competing (but valid) block announcements, which
bifurcates the chain, until one of the two forks is appended to
first. It is also possible and valid for a miner to receive a set of
blocks that retroactively rewrites many blocks; doing so is a
demonstration of computational work that miners accept despite the age
or depth\footnote{The \emph{depth} of a block refers to the number of
  blocks that follow it; the \emph{height} of a block is the number of
  blocks that precede it.} of a rewritten block.

\para{Topology and flooding.} Bitcoin propagates new transaction and
block announcements by flooding throughout a p2p random graph of full
nodes and miners. Each peer in the graph requests direct connections
to 8 other peers, and accepts requests for connections from up to 117
other peers. A peer will offer a newly created transaction to each
neighbor via an \inv message, which reports the hash of the
transaction content as its ID. If a peer does not already possess the
transaction, it will request it using a {\tt getdata} message. Blocks
are handled similarly: \inv messages describe a block by its ID, which
is created from the hash of the block's contents.  Upon receiving the
\inv, peers will request the block if they do not already have it.
Hence, in today's topology, \inv messages cross every edge in the
random graph once, while the actual transaction and block data
typically propagate along only a spanning tree of the graph (more
edges will be traversed if there are propagation delays).  For
convenience, in this paper, we refer to the set of (unconfirmed)
transaction IDs that a peer knows about as the {\em IDpool}. Actual
transaction contents are placed in the mempool.

%\subsection{Bitcoin}
%\para{Accounts.} Bitcoin users store bitcoins in accounts called {\em
%  addresses}, which are created with an empty balance simply by
%generating a public/private key pair. The transfer of coin between
%addresses, via {\em transactions}, is recorded on a public ledger
%called a {\em blockchain}. Transactions are authenticated via a
% private key signature, and the balance of each account can be 
%derived from the blockchain. 
%
%\para{Adding to the blockchain.} To be added to the blockchain,
%transactions are broadcast by users on Bitcoin's p2p network. A set of
%\emph{miners} on the p2p network verify that each transaction is
%signed correctly, does not conflict with a previous transaction, does not
%move more coin than is contained in the address, and other functions.
%Each miner independently agglomerates a set of valid transactions into
%a candidate \emph{block} and attempts to solve a predefined
%cryptographic puzzle as POW, which involves data
%from the candidate block and a specific {\em prior block}. The
%new transactions  are only valid if they do not
%conflict with the set of transactions that are contained in all
%blocks that are direct ancestors.
%
%The first miner to solve the problem broadcasts his solution to the
%network, and by virtue of the solution, is able to add the block to
%the ever-growing blockchain as a child of the prior block. The miners
%then start over, using the newly appended blockchain and the set of
%remaining transactions. The miners' incentive for {\em discovering}
%a new block is a reward of coins, called the {\em coinbase},
%consisting of a predetermined {\em block reward} (currently worth 12.5 Bitcoins)
%and fees from transactions included in the block.
%
%In Bitcoin, the POW computation is dynamically calibrated to take
%approximately ten minutes per block. When transactions appear in a
%block, they are \emph{confirmed}, and each subsequent block
%provides additional confirmation. To announce a new block, a miner
%lists all transactions contained in the new block along with a header
%that contains an easily-verifiable POW solution. When a node or
%miner receives a new block, he validates each transaction in the block
%and the POW.
%
%Notably, if there is a fork on the chain, honest miners always select the prior block as the last block containing the largest amount of POW. However,
%due to propagation delays in the network, 
%miners can receive competing (but valid) block announcements, which
%bifurcates the chain, until one of the two forks is appended to first.
%
%
%\para{Block Headers.} 
%Any entity can elect to be a miner for Bitcoin, and there is no
%centralized party from whom to seek approval for mining. If all miners
%were to simply vote on which block should be appended to the main chain, then the mining process would fall vulnerable to a Sybil
%attack~\cite{Douceur:2002}. The POW puzzle addresses this
%problem by performing a kind of decentralized leader election: the
%miner that solves the puzzle can decide which block to append to the
%chain.
%
%\para{Proof-of-Work.} 
%Bitcoin uses a simple POW algorithm based on cryptographic
%hashing, proposed earlier by Douceur~\cite{Douceur:2002}.
%Specifically, miners apply a 256-bit cryptographic hash algorithm~\cite{hashcash} to
%an 80-byte {\em block header}, and the puzzle is solved if the
%resulting value is less than a known {\em target}, $0<t<2^{256}$. The
%header in Bitcoin consists of the Merkle root of the set of
%transactions, a timestamp, the target (stored as $2^{224}/t$), a {\em
%  nonce}, and the hash of the prior block's header. If the hash is not
%less than the target, then a new nonce is selected to generate a new
%hash (the Merkle root can be adjusted as well). This process repeats
%until some miner finds a solution.
%  
%Each time a nonce is selected and the block
%header is hashed, the miner is sampling a value from a discrete uniform
%distribution with range $[0,2^{256}-1]$. The probability of solving the POW and
%discovering a block is the cumulative probability of selecting a value
%from $[0,t]$, which is $t/{2^{256}}$. Hence, in expectation, the
%number of samples needed to discover a block is ${2^{256}/t}$. Bitcoin
%adjusts the target so that on average it takes about 600 seconds to
%find a block. Typically, the target is described for convenience as a
%{\em difficulty}, defined to be $D=2^{224}/t$. Bitcoin's difficulty
%is set once every two weeks.

